### 2016-06-06 coproject/figures/linear.tex
UPD: Added some post processing and written some part of the report

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=4.585417in,
height=3.591094in,
at={(0.769167in,0.484687in)},
scale only axis,
axis on top,
xmin=0.5,
xmax=10.5,
xtick={1,2,3,4,5,6,7,8,9,10},
xticklabels={{0},{1},{2},{3},{4},{5},{6},{7},{8},{9}},
xlabel={Truth},
y dir=reverse,
ymin=0.5,
ymax=10.5,
ytick={1,2,3,4,5,6,7,8,9,10},
yticklabels={{0},{1},{2},{3},{4},{5},{6},{7},{8},{9}},
ylabel={Prediction}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=10.5,ymin=0.5,ymax=10.5] {linear-1.png};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:1,1,0) {968};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,3,0) {10};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,4,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,5,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,6,0) {8};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,7,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,8,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,9,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:1,10,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,1,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:2,2,0) {1124};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,3,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,4,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,5,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,6,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,7,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,8,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:2,10,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,2,0) {2};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:3,3,0) {970};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,4,0) {17};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,5,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,6,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,7,0) {9};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,8,0) {21};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:3,10,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,2,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,3,0) {9};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:4,4,0) {952};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,5,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,6,0) {32};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,7,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,8,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,9,0) {23};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:4,10,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,3,0) {6};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,4,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:5,5,0) {942};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,6,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,7,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,8,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,9,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:5,10,0) {29};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,1,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,2,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,3,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,4,0) {14};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,5,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:6,6,0) {807};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,7,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,8,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,9,0) {28};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:6,10,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,2,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,3,0) {9};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,4,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,5,0) {7};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,6,0) {9};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:7,7,0) {923};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,8,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,9,0) {6};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:7,10,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,1,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,2,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,3,0) {8};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,4,0) {12};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,5,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,6,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,7,0) {0};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:8,8,0) {966};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,9,0) {5};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:8,10,0) {15};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,1,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,2,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,3,0) {14};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,4,0) {13};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,5,0) {2};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,6,0) {19};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,7,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,8,0) {1};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:9,9,0) {888};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:9,10,0) {4};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,1,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,2,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,3,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,4,0) {1};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,5,0) {23};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,6,0) {3};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,7,0) {0};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,8,0) {19};
\node[align=center, inner sep=0mm, text=black]
at (axis cs:10,9,0) {4};
\node[align=center, inner sep=0mm, text=white]
at (axis cs:10,10,0) {931};
\end{axis}
\end{tikzpicture}%