### 2016-06-06 coproject/figures/train-set.tex

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-1.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-2.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-3.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-4.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-5.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-6.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 1}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-7.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-8.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-9.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-10.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 9}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-11.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-12.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 4}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-13.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-14.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={label: 8}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/train-set/train-set-15.png};
\end{axis}
\end{tikzpicture}%