### 2016-06-07 coproject/figures/wrong-label.tex
UPD: Finished writing experience + conclusion. Just need to review the whole report

% This file was created by matlab2tikz.
% Minimal pgfplots version: 1.3
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\begin{tikzpicture}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-1.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={5 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-2.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-3.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={3 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-4.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={2 $\rightarrow$ 0}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-5.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={7 $\rightarrow$ 9}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-6.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(8.490105cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-7.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={5 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-8.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-9.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 6}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-10.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(6.367579cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={7 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-11.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(0cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={2 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-12.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(4.245052cm,0cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={4 $\rightarrow$ 2}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-13.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,5.468586cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={9 $\rightarrow$ 3}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-14.png};
\end{axis}

\begin{axis}[%
width=1.509895cm,
height=1.41923cm,
at={(2.122526cm,2.734293cm)},
scale only axis,
axis on top,
xmin=0.5,
xmax=28.5,
tick align=outside,
y dir=reverse,
ymin=0.5,
ymax=28.5,
hide axis,
title style={},
title={3 $\rightarrow$ 7}
]
\addplot [forget plot] graphics [xmin=0.5,xmax=28.5,ymin=0.5,ymax=28.5] {./figures/wrong-label/wrong-label-15.png};
\end{axis}
\end{tikzpicture}%